---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
head(data)
```

## What is mean total number of steps taken per day?
```{r}
data_agg_steps <- aggregate(steps ~ date, data, sum)
head(data_agg_steps)
histo <- hist(data_agg_steps$steps, breaks = 10)
meansteps <- mean(data_agg_steps$steps)
mediansteps <- median(data_agg_steps$steps)
print(paste(meansteps, mediansteps))
```

The mean number of steps taken per day is `r format(meansteps, scientific=FALSE)`. The median number of steps taken per day is `r mediansteps`.

## What is the average daily activity pattern?
```{r}
data_agg_interval <- aggregate(steps ~ interval, data, mean)
plotline <- plot(data_agg_interval, type='l')
maxsteps_interval <- data_agg_interval[which(data_agg_interval$steps == max(data_agg_interval$steps)), 1]
print(maxsteps_interval)
```
The 5-minute interval labeled `r maxsteps_interval` on average has the maximum number of steps.

## Inputing missing values
```{r}
num_na <- sum(is.na(data))
print(num_na)
```
The number of missing values is `r num_na`. Filling missing values with the average for that interval. 
```{r}
for (ii in 1:length(data_agg_interval[[1]])) {
     data[(is.na(data)) & (data$interval ==  data_agg_interval$interval[[ii]]),]$steps <- data_agg_interval$steps[[ii]]
}
head(data)
data_agg_steps_fill <- aggregate(steps ~ date, data, sum)
histo_fill <- hist(data_agg_steps_fill$steps, breaks=10)
meansteps_fill <- mean(data_agg_steps_fill$steps)
mediansteps_fill <- median(data_agg_steps_fill$steps)
print(paste(meansteps_fill, mediansteps_fill))
```

The mean number of steps taken per day from filled in data is `r format(meansteps_fill, scientific=FALSE)`. The median number of steps taken per day from filled in data is `r format(mediansteps_fill, scientific=FALSE)`. 

The mean remains the same from earlier but the median changes to match the mean with this method of filling missing values.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data[['day']] <- weekdays(as.Date(data$date))
data[['weekday']] <- ifelse(data$day == 'Saturday' | data$day == 'Sunday', 1, 0)
data_avg_steps_weekday <- aggregate(steps ~ interval, data[data$weekday==0,], mean)
data_avg_steps_weekend <- aggregate(steps ~ interval, data[data$weekday==1,], mean)
par(mfrow=c(2,1))
plot(data_avg_steps_weekday, type='l', main='Weekday', ylim=c(-25, 250))
plot(data_avg_steps_weekend, type='l', main='Weekend', ylim=c(-25, 250))
```
